### 2015-01-19 report/timedomain.tex

The equations of consideration write as
 \begin{align}
 \epsilon_0\partial_t E_{x}-\partial_y H_z=-eN_e u_x, \nonumber\\
 \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
 \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
 m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
 m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
 \end{align}

We compute a numerical approximation of~\eqref{eq:tdpb} on a Cartesian grid using the Yee scheme. Let $n$ denote the time step, as usual $t_n = n \Delta t$. $H_z$ will then be taken on a half time step, and our equation, discretized in them write :
 \begin{align}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}-\partial_y H_z=-eN_e u_x, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z=-eN_e u_y,\nonumber\\
 \mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y-\partial_y E_x=0, \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
 \end{align}
 We discretize our equation on staggered grid, meaning $E_x$, $u_x$, $E_y$, $u_y$ and $H_z$ are not and the same grid. Evidently we already stated it for $H_z$ in the time discretization. Fig.~\ref{schemepos} shows the positions of $Ex$, $u_x$ (black dots), $E_y$ $u_y$ (blue dots) and $H_z$ (red dots) on the discretized grid in space and time. We obtain the following discretized equations,

\be 
\eps_0 \frac{E_x\inp- E_x\midin}{\Delta t} = -e N_e \frac{u_x\inp + u_x \midin}{2},
\label{eq:ns1}
\ee 
\be
\eps_0 \frac{E_y\ihnp-E_y\ihn}{\Delta t}+ \frac{H_z\ipnh-H_z\inh}{\Delta x} = -eN_x \frac{u_y\ihnp + u_y\ihn}{2} ,
\label{eq:ns2}
\ee
\be
 \frac{H_z\inh-H_z\inmh}{\Delta t}+ \frac{E_y\ihn-E_y\imhn}{\Delta x}=0,
 \label{eq:ns3}
\ee
\be 
m_e \frac{u_x\inp-u_x\midin }{\Delta x} = e N_e \frac{E_x\inp+E_x\midin}{2}-\nu m_e\frac{u_x\inp+u_x \midin}{2}+ eB_0\frac{u_y\ihn+u_y\ihnp}{2},
\label{eq:ns4}
\ee
\be
m_e \frac{u_y\ihnp-u_y\ihn}{\Delta t} = eN_e \frac{E_y\ihnp + E_y\ihn}{2}- \nu m_e \frac{u_y\ihnp +u_y\ihn}{2}-eB_0 \frac{u_x\midin+u_x\inp}{2},
\ee
\input{grid}

We first compute $H_z$ at time $n+1/2$ with \eqref{eq:ns3}. Then, \eqref{eq:ns2} gives the expression $E_y\ihnp$ as a function of known variables, and $u_y^{n+1}$, as follow
\be
E_y\ihnp = E_y\ihn - \frac{\Delta t }{\eps_0}\frac{H_z\ipnh-H_z\inh}{\Delta x} - \frac{\Delta t e N_e}{\eps_0}\frac{u_y\ihnp + u_y\ihn}{2}.
\label{eq:ns5}
\ee

By incorporating this expression into \eqref{eq:ns5} we then obtain $u_y$ at time $n+1$ as a function of known variables and $u_x^{n+1}$, 
\be 
\begin{array}{l}
u_y\ihnp m_e = m_e u_y\ihn + \Delta t eN_e \frac{E_y\ihn}{2}\\
+ \Delta teN_e\left(E_y\ihn - \frac{\Delta t }{\eps_0}\frac{H_z\ipnh-H_z\inh}{\Delta x} - \frac{\Delta t e N_e}{\eps_0}\frac{u_y\ihnp + u_y\ihn}{2}\right) \\
-\nu \Delta t m_e \frac{u_y\ihnp +u_y\ihn}{2}-\Delta teB_0 \frac{u_x\midin-u_x\inp}{2},
\end{array}
\label{eq:ns6}
\ee
which gives the following expression for $u_y\ihnp$

\begin{align}
u_y\ihnp \left(m_e(1+\frac{\Delta t \nu}{2}) +\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}\right) =&\left(m_e(1-\frac{\Delta t \nu}{2}) -\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}\right)u_y\ihn +\Delta t eN_e E_y\ihn \nonumber\\
&- \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x} - \frac{e Bo \Delta t}{2}\left(u_x\midin+u_x\inp\right),
\label{eq:ns7} 
\end{align}
which we write 
\begin{align}
u_y\ihnp K_1&= K_2 u_y\ihn +\Delta t eN_e E_y\ihn-  \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x}- \frac{e Bo \Delta t}{2}\left(u_x\midin+u_x\inp\right).
\label{eq:ns7bis} 
\end{align}
with $K_1 =  m_e(1+\frac{\Delta t \nu}{2}) +\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}$ and $K_2 =m_e(1- \frac{\Delta t \nu}{2}) -\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}$.

Now we just need an expression of $E_x\inp$ as a function of time $n$ variables and $u_x^{n+1}$ and putting \eqref{eq:ns7} in \eqref{eq:ns4}, the scheme will be explicit. 
We can find such and expression in \eqref{eq:ns1}
\be 
E_x\inp = E_x\midin - \frac{\Delta t e N_e}{\eps_0}\frac{u_x\inp + u_x\midin}{2}.
\label{eq:ns8}
\ee

We then have the scheme

\begin{align}
m_e u_x\inp= & m_e u_x\midin  + \Delta t e N_e \frac{1}{2}\left( E_x\midin - \frac{\Delta t e N_e}{\eps_0}\frac{u_x\inp + u_x\midin}{2}\right) + \Delta t  e N_e \frac{1}{2}E_x\midin -\Delta t\nu m_e\frac{u_x\inp+u_x \midin}{2}\nonumber\\
&+  \frac{e B_O \Delta t}{2K_1}\Big(K_2 u_y\ihn +\Delta t eN_e E_y\ihn -  \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x}\nonumber\\
&- \frac{e Bo \Delta t}{2}\left(u_x\midin+u_x\inp\right)\Big)+  \frac{1}{2}\Delta teB_0u_y\ihn,
\label{eq:ns9}
\end{align}

which one can simplify into
\begin{align}
u_x & \left(me(1+\frac{\Delta t \nu}{2})+ \frac{(eB_0\Delta t)^2}{4K_1} + \frac{(e\Delta t N_e)^2}{4\eps_0}\right) u_x\inp 
= \left(me(1-\frac{\Delta t \nu}{2})- \frac{(eB_0\Delta t)^2}{4K_1} -\frac{(e\Delta t N_e)^2}{4\eps_0}\right) u_x\midin \nonumber\\
& + e \Delta t N_e E_x \midin  + \left( \frac{eB_0 \Delta t K_2}{2 K_1}+ \frac{e B_0 \Delta t}{2}\right) u_y\ihn - \frac{(e\Delta t)^2 N_e B_0}{2K_1}E_y\ihn  - \frac{e^2(\Delta t)^3 B_0}{4K_1}\frac{H_z \ipnh- H_z \inh}{\Delta x},
&\label{eq:ns10} 
\end{align}
and the scheme is then fully implicit.