### 2015-01-20 report/proc.tex
adding first batch of refs

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for Proccedings 
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%%   EDP Sciences  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[proc]{edpsmath}
%
%\usepackage{amsfonts,graphicx,amssymb}%amsfonts pour les \mathbb
%\usepackage{amsmath,amsthm,amssymb}
%%%%%%%%%%%%%--PREAMBLE--%%%%%%%%%%%%%%%%%%
%%-----------------------------
%%         ...........
%%         your macros
\usepackage[frenchb,english]{babel}
\usepackage{macros}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}
%\usepackage{diagbox}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{red}{rgb}{1,0.3,0.3}
\usetikzlibrary{positioning,shapes,shadows,arrows}

%%         ...........
%%-------------------------%%----
%%%%%%%%%%%%%%%--BODY--%%%%%%%%%%%%%%%%%%

\begin{document}

%%-----------------------------
%%      the top matter
%%-----------------------------
\title{Study of the
	numerical solution of X-Modes equations around the hybrid resonnance} %\thanks{...}\thanks{...}% At most 5 thanks
%
\author{C\'eline Caldini-Queiros}\address{Max Planck Institute für PlasmaPhysik, Garching bei Muenchen}
\author{Bruno Depsr\'es}\address{Laboratoire Jacques-Louis Lions, University PARIS 6}
\author{Lise-Marie Imbert-Gérard}\address{Courant Institute of Mathematical Sciences, New York University}
\author{Maryna Kachanovska}\address{POEMS, INRIA, ENSTA ParisTech, Paris}

%
%\dedicated{\it Dedicated to Maurice Dupont} %if necessary
%
\begin{abstract} ... \end{abstract}
%
\begin{resume} ... \end{resume}
%
%
\maketitle
\textcolor{red}{should we add Olivier Lafitte and Remi Sart anywhere in the proceeding ?}\\
%!!!!!!!!!!!todo !!!!!!!!!!!

%\item !!!!to do : correct proof of coercivity (celine)
%\item !!!!to do : write Yee scheme 			(celine)
%
%%with explicit in \nu coercivity and continuity constants

\input{intro}
\input{equations}
\input{numerics}
\input{timedomain}
%\input{absorption}
\begin{thebibliography}{99}
\bibitem{stable_yee_plasma_current} F. da Silva M. Campos Pinto, B. Despres and S. Heuraux, Stable Coupling of the Yee Scheme with a Linear Current Model,
2014
\bibitem{Depres_2014} Després, B., Imbert-Gérard, L. M. and Weder, R. (2014). Hybrid resonance of Maxwell's equations in slab geometry. Journal de Mathématiques Pures et Appliquées, 101(5), 623-659.
\bibitem{LMIG_thesis} Imbert-Gérard, L. M. (2013). Analyse mathématique et numérique de problèmes d'ondes apparaissant dans les plasmas magnétiques (Doctoral dissertation, Université Pierre et Marie Curie-Paris VI).
\bibitem{brenner} Brenner, S. C. and Scott, R. (2008). The mathematical theory of finite element methods (Vol. 15). Springer.
\bibitem{Stix} Stix, T. H. (1992). Waves in plasmas. Springer.
\end{thebibliography}

\end{document}