### 2015-01-26 report/abstract.tex
english abstract : first rough draft

Hybrid resonnance is a physical phenomenon that appears for example in the heating of plasma, and as such is of interest in the developping of new energy source, like for instance in the ITER project. In this paper we focus on the  non-regularity of
the solutions of Maxwell equations in plasmas under strong background magnetic field. Our purpose here is two-fold. On one hand we investigate the finite element approximation of the one dimensionnal problem written in the frequency domain, and on the other hand we investigate the finite difference approximation of the one dimensionnal time dependant problem. We will also compare the results of theses to differents methods.