### 2015-01-27 report/abstract.tex
corrected abstract

Hybrid resonance is a physical phenomenon 
that appears for example in the heating of plasma, 
and as such is of interest in the developing of new sources of energy, 
e.g. in the ITER project. In this paper we focus on the non-regularity of
the solutions of Maxwell equations in plasmas 
under strong background magnetic field. 
Our purpose here is two-fold. 
On one hand we investigate the finite element approximation of the one-dimensional problem 
written in the frequency domain, 
and on the other hand we study the finite difference approximation 
of the one-dimensional time dependant problem. 
We compare the results of these two differents methods in the context of limiting absorption and limiting amplitude principles.