### 2015-01-27 report/proc.tex

\documentclass[proc]{edpsmath}

\usepackage{macros}

% intro 
% theoretical part 
% numerical part : frequential domain
% numerical part : time dependant scheme
% numerical part : comparison
% concl.

\begin{document}
\begin{itemize}
\item !!!!to do : correct proof of coercivity (celine)
\item !!!!to do : write Yee scheme 			(celine)
\item !!!!to do : well posedness = lemma		(celine)
%with explicit in \nu coercivity and continuity constants 
\end{itemize}

\input{intro}
\input{equations}
\input{numerics}
\begin{thebibliography}{99}
\bibitem{stable_yee_plasma_current} F. da Silva, M. Campos Pinto, B. Despres, S. Heuraux, Stable Coupling of the Yee Scheme with a Linear Current Model,
2014
\end{thebibliography}

\end{document}