### 2015-01-28 report/proc.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for Proccedings 
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%%   EDP Sciences  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[proc]{edpsmath}
%
%\usepackage{amsfonts,graphicx,amssymb}%amsfonts pour les \mathbb
%\usepackage{amsmath,amsthm,amssymb}
%%%%%%%%%%%%%--PREAMBLE--%%%%%%%%%%%%%%%%%%
%%-----------------------------
%%         ...........
%%         your macros
\usepackage[frenchb,english]{babel}
\usepackage{macros}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}
\usepackage{diagbox}
\usepackage{color}
\usepackage{placeins}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{df}{Definition}[section]
\newtheorem{lemma}[theorem]{Lemma}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{red}{rgb}{1,0.3,0.3}
%\usetikzlibrary{positioning,shapes,shadows,arrows}

%%         ...........
%%-------------------------%%----
%%%%%%%%%%%%%%%--BODY--%%%%%%%%%%%%%%%%%%

\begin{document}
\newcommand{\urev}[1]{{\color{red}#1}}
\newcommand{\mrev}[1]{{\color{blue}#1}}

%%-----------------------------
%%      the top matter
%%-----------------------------
\title{Study of the
	numerical solution of X-Modes equations around the hybrid resonance} %\thanks{...}\thanks{...}% At most 5 thanks
%
\author{C\'eline Caldini-Queiros}\address{Max Planck Institute für PlasmaPhysik, Garching bei Muenchen}
\author{Bruno Despr\'es}\address{Laboratoire Jacques-Louis Lions, University PARIS 6}
\author{Lise-Marie Imbert-Gérard}\address{Courant Institute of Mathematical Sciences, New York University}
\author{Maryna Kachanovska}\address{POEMS, INRIA, ENSTA ParisTech, Paris}

%
%\dedicated{\it Dedicated to Maurice Dupont} %if necessary
%
\begin{abstract}
\input{abstract}
\end{abstract}
%
\begin{resume} 
\input{resume}
\end{resume}
%
%
\maketitle
\textcolor{red}{should we add Olivier Lafitte and Remi Sart anywhere in the proceeding ?}\\

\input{intro}
\input{FD}%\input{equations}

\input{timedomain}
%\input{freqdomain}
\input{numerics}
\input{absorption}
\input{conclusions}

\begin{thebibliography}{99}
\bibitem{abramowitz_stegun} Abramowitz M. and Stegun I. A., eds. (1972). Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables. New York: Dover Publications.	Handbook of mathematical functions
\bibitem{brenner} Brenner, S. C. and Scott, R. (2008). The mathematical theory of finite element methods (Vol. 15). Springer.
\bibitem{stable_yee_plasma_current} Da Silva, F., Campos-Pinto, M., Després, B., and Heuraux, S. (2014). Stable coupling of the Yee scheme with a linear current model.
\bibitem{reflectometers_2010}Da Silva, F., Heuraux, S., Gusakov, E. Z., and Popov, A. (2010). A Numerical Study of Forward-and Backscattering Signatures on Doppler-Reflectometry Signals. Plasma Science, IEEE Transactions on, 38(9), 2144-2149.
\bibitem{reflectometers_2006} Da Silva, F., Heuraux, S and  Manso, M. (2006). Developments on reflectometry simulations for fusion plasmas: application to ITER position reflectometers. Journal of plasma physics, 72(06), 1205-1208.
\bibitem{Despres_2014} Després, B., Imbert-Gérard, L. M. and Weder, R. (2014). Hybrid resonance of Maxwell's equations in slab geometry. Journal de Mathématiques Pures et Appliquées, 101(5), 623-659.
\bibitem{Dumont_2005} Dumont, R. J., Phillips, C. K., and Smithe, D. N. (2005). Effects of non-Maxwellian species on ion cyclotron waves propagation and absorption in magnetically confined plasmas. Physics of Plasmas (1994-present), 12(4), 042508.
\bibitem{Freidberg_2007} Freidberg, J. P. (2007). Plasma physics and fusion energy. Cambridge university press.
\bibitem{LMIG_thesis} Imbert-Gérard, L. M. (2013). Analyse mathématique et numérique de problèmes d'ondes apparaissant dans les plasmas magnétiques (Doctoral dissertation, Université Pierre et Marie Curie-Paris VI).
\bibitem{Morawetz} Morawetz, C. S. (1962). The limiting amplitude principle. Communications on Pure and Applied Mathematics, 15(3), 349-361.
\bibitem{Mouhot_2011} Mouhot, C., and Villani, C. (2011). On landau damping. Acta mathematica, 207(1), 29-201.
\bibitem{Stix} Stix, T. H. (1992). Waves in plasmas. Springer.
 \bibitem{Xu_2006}Xu, L., ans Yuan, N. (2006). FDTD formulations for scattering from 3-D anisotropic magnetized plasma objects. Antennas and Wireless Propagation Letters, IEEE, 5(1), 335-338.
\end{thebibliography}

\end{document}