### 2015-01-29 report/proc.tex
Bruno's proc.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for Proccedings 
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%%   EDP Sciences  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[proc]{edpsmath}
%
%\usepackage{amsfonts,graphicx,amssymb}%amsfonts pour les \mathbb
%\usepackage{amsmath,amsthm,amssymb}
%%%%%%%%%%%%%--PREAMBLE--%%%%%%%%%%%%%%%%%%
%%-----------------------------
%%         ...........
%%         your macros
\usepackage[frenchb,english]{babel}
\usepackage{macros}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}
\usepackage{diagbox}
\usepackage{color}
\usepackage{placeins}
%\newenvironment{remark}[1][Remark]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{df}{Definition}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{red}{rgb}{1,0.3,0.3}
%\usetikzlibrary{positioning,shapes,shadows,arrows}

%%         ...........
%%-------------------------%%----
%%%%%%%%%%%%%%%--BODY--%%%%%%%%%%%%%%%%%%

\begin{document}
\newcommand{\urev}[1]{{\color{red}#1}}
\newcommand{\mrev}[1]{{\color{blue}#1}}

%%-----------------------------
%%      the top matter
%%-----------------------------
\title{A numerical study of the
	solution of X-Mode equations around the hybrid resonance}
\thanks{\rm The authors acknowledge 
	the support
	of ANR under contract ANR-12-BS01-0006-01. Moreover 
	this work was carried out within the framework of the European Fusion Development Agreement and 
	the French Research Federation for Fusion Studies. 
	It is supported by the European Communities under the contract of Association between Euratom and CEA. 
	The views and opinions expressed herein do not necessarily reflect those of the European Commission.} %\thanks{...}\thanks{...}% At most 5 thanks
%
\author{C\'eline Caldini-Queiros}\address{Max Planck Institute für PlasmaPhysik, Garching bei Muenchen}
\author{Bruno Despr\'es}\address{Laboratoire Jacques-Louis Lions, University PARIS 6}
\author{Lise-Marie Imbert-Gérard}\address{Courant Institute of Mathematical Sciences, New York University}
\author{Maryna Kachanovska}\address{POEMS, INRIA, ENSTA ParisTech, Paris}

%
%\dedicated{\it Dedicated to Maurice Dupont} %if necessary
%
\begin{abstract}
\input{abstract}
\end{abstract}
%
\begin{resume} 
\input{resume}
\end{resume}
%
%
\maketitle
%\textcolor{red}{should we add Olivier Lafitte and Remi Sart anywhere in the proceeding ?}\\

\input{intro}
\input{FD}%\input{equations}

\input{timedomain}
%\input{freqdomain}
\input{numerics}
\input{absorption}
\input{conclusions}

\begin{thebibliography}{99}
\bibitem{cedar}
Charles, F., Despr\'es, B., Merhenberg, M, 
Enhanced convergence estimates for semi-lagrangian schemes Application to the Vlasov-Poisson equation, SIAM J. Numer. Anal. 2013, 51(2), 840-863.

\bibitem{stable_yee_plasma_current} Da Silva, F., Campos-Pinto, M., Després, B., and Heuraux, S. (2014). Stable coupling of the Yee scheme with a linear current model.

\bibitem{compfluids}
S. Del Pino, B. Despr\'es, P. Hav\'e, H. Jourdren, and P. F. Piserchia. 3D finite
volume simulation of acoustic waves in the earth atmosphere.
Comput. \&
Fluids, 38(4):765-777, 2009.

\bibitem{reflectometers_2006} Da Silva, F., Heuraux, S and  Manso, M. (2006). Developments on reflectometry simulations for fusion plasmas: application to ITER position reflectometers. Journal of plasma physics, 72(06), 1205-1208.

\bibitem{reflectometers_2010}Da Silva, F., Heuraux, S., Gusakov, E. Z., and Popov, A. (2010). A Numerical Study of Forward-and Backscattering Signatures on Doppler-Reflectometry Signals. Plasma Science, IEEE Transactions on, 38(9), 2144-2149.

\bibitem{Despres_2014} Després, B., Imbert-Gérard, L. M. and Weder, R. (2014). Hybrid resonance of Maxwell's equations in slab geometry. Journal de Mathématiques Pures et Appliquées, 101(5), 623-659.

\bibitem{Dumont_2005} Dumont, R. J., Phillips, C. K., and Smithe, D. N. (2005). Effects of non-Maxwellian species on ion cyclotron waves propagation and absorption in magnetically confined plasmas. Physics of Plasmas (1994-present), 12(4), 042508.

\bibitem{Freidberg_2007} Freidberg, J. P. (2007). Plasma physics and fusion energy. Cambridge university press.

\bibitem{phdhattori}
T. Hattori, T. (2014).
D\'ecomposition de domaine pour la simulation Full-Wave dans un plasma froid.
Ph.D. Thesis, Universit\'e de
Lorraine.

\bibitem{LMIG_thesis} Imbert-Gérard, L. M. (2013). Analyse mathématique et numérique de problèmes d'ondes apparaissant dans les plasmas magnétiques (Doctoral dissertation, Université Pierre et Marie Curie-Paris VI).

\bibitem{brenner} Brenner, S. C. and Scott, R. (2008). The mathematical theory of finite element methods (Vol. 15). Springer.

\bibitem{simon}
Simon Labrunie, Pierre Bertrand, Jean-Rodolphe Roche, Aurore Back,
Takashi Hattori, 
Electromagnetic wave propagation and absorption in
magnetised plasmas: variational formulations and
domain decomposition, Hal preprint 2014, https://hal.archives-ouvertes.fr/hal-01075137.

\bibitem{Stix} Stix, T. H. (1992). Waves in plasmas. Springer.

\bibitem{Morawetz} Morawetz, C. S. (1962). The limiting amplitude principle. Communications on Pure and Applied Mathematics, 15(3), 349-361.

\bibitem{Mouhot_2011} Mouhot, C., and Villani, C. (2011). On Landau damping. Acta mathematica, 207(1), 29-201.

\bibitem{abramowitz_stegun} Abramowitz M. and Stegun I. A., eds. (1972). Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables. New York: Dover Publications.
Handbook of mathematical functions
\end{thebibliography}

\end{document}