### 2015-01-29 report/abstract.tex
Bruno's abstract/resume/intro/FD/timedomain

Hybrid resonance is a physical phenomenon
that appears for example in the heating of plasma,
and as such is of scientific interest in the development  of  the ITER project.
In this paper we focus some solutions with low regularity of Maxwell equations in
plasmas under strong background magnetic field.
Our purpose is two-fold.
On one hand we investigate the finite element approximation of the one dimensional problem 
written in the frequency domain, and on the other hand we investigate two different 
finite difference approximations of the one dimensional time dependent problem.
We will also compare the results of these different methods.