### 2015-02-04 report_clean/figure_generator/pics_frequency_domain/res_sol.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[frenchb,english]{babel}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}

\usepackage{color}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{red}{rgb}{1,0.3,0.3}
\usetikzlibrary{positioning,shapes,shadows,arrows}
\begin{document}

\begin{figure}
\pagenumbering{gobble}
\begin{tikzpicture}
    \begin{axis}[
    	width=\textwidth,
    	height=\textwidth,
    	xmin=-15,
    	xmax=10,
        xlabel=$x$,
        ylabel=$E_{x,y}$,
        legend style={
at={(0.2,1.5)},
legend pos= south east,
anchor=south east, 
draw=none, 
font=\small
}
]
\addplot[blue] table {pics_resonance/ime1.0.1.dat}; 
\addlegendentry{$\Im{E_x^{\nu}},\nu=10^{-1}$};
\addplot[red] table {pics_resonance/ime1.0.01.dat}; 
\addlegendentry{$\Im{E_x^{\nu}},\nu=10^{-2}$};
\addplot[black] table {pics_resonance/ime1.0.001.dat}; 
\addlegendentry{$\Im{E_x^{\nu}},\nu=10^{-3}$};
\addplot[blue,dashed] table {pics_resonance/ree2.0.1.dat}; 
\addlegendentry{$\Re{E_y^{\nu}},\nu=10^{-1}$};
\addplot[red,dashed] table {pics_resonance/ree2.0.01.dat}; 
\addlegendentry{$\Im{E_y^{\nu}},\nu=10^{-2}$};
\addplot[black,dashed] table {pics_resonance/ree2.0.001.dat}; 
\addlegendentry{$\Im{E_y^{\nu}},\nu=10^{-3}$};
 \end{axis}
\end{tikzpicture}
\end{figure}

\end{document}