### 2015-02-04 report_clean/figure_generator/pics_time_domain/airy/figure_nu1e4.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=1.2\textwidth, 
  height=0.5\textwidth,
   xlabel={$t$},
  enlarge x limits=false,
  ymax=0.4,
  xtick={2.3976e5,2.3982e5,2.3988e5,2.3994e5, 2.39999e5},
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_y^{\nu}(x_c,t)$},
  legend style={draw=none, fill=white, font=\tiny}
   ]
 \addplot[color=blue] file{airy/td_x_0.45_nu0.0001.dat};
  \addlegendentry{$E_y^{\nu}(x_c,t),\; x_c=0.45, \; \nu=10^{-4}$};
  \addplot[color=red, mark=*, dashed, mark size=0.6] file{airy/fd_x_0.45_nu0.0001.dat};
  \addlegendentry{$\Im\left(\hat{E}_y^{\nu}(x_c)\mathrm{e}^{it}\right),\; x_c=0.45, \; \nu=10^{-4}$};
  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}