### 2015-02-04 report_clean/figure_generator/pics_time_domain/res/ex_fixed_t.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tabular}{c}
\begin{tikzpicture}
 \begin{axis}[
  width=\textwidth, 
  height=0.5\textwidth,
  xlabel={$x$},
  enlarge x limits=false,
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_x^{\nu}(x,t_c)$},
  legend style={draw=none, fill=white, font=\small}
   ]
     \addplot[color=red, dashed] file{nu_1e2/E1_fd_t_29765.dat};
  \addlegendentry{$\Im\left(\hat{E}_x^{\nu}(x)\mathrm{e}^{it_c}\right),\; t_c=29765, \; \nu=10^{-2}$};
 \addplot[color=blue] file{nu_1e2/E1_td_t_29765.dat};
  \addlegendentry{$E_x^{\nu}(x,t_c),\; t_c=29765, \; \nu=10^{-2}$};
  \end{axis}
\end{tikzpicture}\\
\begin{tikzpicture}
 \begin{axis}[
  width=\textwidth, 
  height=0.5\textwidth,
  xlabel={$x$},
  enlarge x limits=false,
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_x^{\nu}(x,t_c)$},
  legend style={draw=none, fill=white, font=\small}, 
  legend pos=south east
   ]
 \addplot[color=red, dashed] file{nu_1e2/E1_fd_t_29800.dat};
 \addlegendentry{$\Im\left(\hat{E}_x^{\nu}(x)\mathrm{e}^{it_c}\right),\; t_c=29800, \; \nu=10^{-2}$};

 \addplot[color=blue] file{nu_1e2/E1_td_t_29800.dat};
 \addlegendentry{$E_x^{\nu}(x,t_c),\; t_c=29800, \; \nu=10^{-2}$};
  \end{axis}
\end{tikzpicture}
\end{tabular}
\end{figure}
\end{document}