### 2015-02-04 report_clean/figure_generator/pics_time_domain/res/ex_fixed_x_nu1e4.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=1.2\textwidth, 
  height=0.5\textwidth,
   xlabel={$t$},
  enlarge x limits=false,
  ymax=2,
  xmin=1.1988e5,
  xtick={1.1988e5,1.1994e5,1.1999e5},
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_x^{\nu}(x_c,t)$},
  legend style={draw=none, fill=white, font=\tiny}, 
  legend columns=2
   ]

   \addplot[color=blue,each nth point=10] file{nu_1e4/e1_td_point0.125.dat};
  \addlegendentry{$E_x^{\nu}(x_c,t),\; x_c=0.0125, \; \nu=10^{-4}$};
    \addplot[color=red, dashed, each nth point=10] file{nu_1e4/e1_fd_point0.125.dat};
  \addlegendentry{$\Im\left(\hat{E}_x^{\nu}(x_c)\mathrm{e}^{it}\right),\; x_c=0.0125, \; \nu=10^{-4}$};

   \addplot[color=green,each nth point=10] file{nu_1e4/e1_td_point2.5.dat};
  \addlegendentry{$E_x^{\nu}(x_c,t),\; x_c=-2.5, \; \nu=10^{-4}$};
      \addplot[color=black, dashed, each nth point=10] file{nu_1e4/e1_fd_point2.5.dat};
  \addlegendentry{$\Im\left(\hat{E}_x^{\nu}(x_c)\mathrm{e}^{it}\right),\; x_c=-2.5, \; \nu=10^{-4}$};
  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}