### 2015-02-04 report_clean/figure_generator/pics_time_domain/res/ey_fixed_t_1e3.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=\textwidth, 
  height=0.5\textwidth,
  xlabel={$x$},
  ymax=0.6,
  enlarge x limits=false,
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_y^{\nu}(x,t_c)$},
  legend style={draw=none, fill=white, font=\small}
   ]

 \addplot[color=blue] file{nu1e3/Ey_nu1e3_t_29765_TD.dat};
  \addlegendentry{$E_y^{\nu}(x,t_c),\; t_c=29765, \; \nu=10^{-3}$};
     \addplot[color=green, dashed, each nth point=2] file{nu1e3/Ey_nu1e3_t_29765_FD.dat};
  \addlegendentry{$\Im\left({\hat{E}_y^{\nu}(x_c)\mathrm{e}^{it}}\right),\; t_c=29765, \; \nu=10^{-3}$};
  \addplot[color=red, each nth point=2] file{nu1e3/Ey_nu1e3_t_29800_FD.dat};
  \addlegendentry{$\Im\left({\hat{E}_y^{\nu}(x_c)\mathrm{e}^{it}}\right),\; t_c=29800, \; \nu=10^{-3}$};

  \addplot[color=black, dashed] file{nu1e3/Ey_nu1e3_t_29800_TD.dat};
  \addlegendentry{$E_y^{\nu}(x,t_c),\; t_c=29800, \; \nu=10^{-3}$};
  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}