### 2015-02-04 report_clean/figure_generator/pics_time_domain/res/ey_fixed_x_1e3.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=1.2\textwidth, 
  height=0.5\textwidth,
   xlabel={$t$},
  enlarge x limits=false,
  ymax=0.5,
  xmin=2.9955e4,
  %xtick={2.97e4,2.98e4,2.99e4,3e4},
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_y^{\nu}(x_c,t)$},
  legend style={draw=none, fill=white, font=\tiny}
   ]
      \addplot[color=blue] file{nu1e3/Ey_nu1e3_x2.5125TD.dat};
  \addlegendentry{$E_y^{\nu}(x_c,t),\; x_c=-2.5125, \; \nu=10^{-3}$};
  \addplot[color=red, dashed] file{nu1e3/Ey_nu1e3_x2.5125FD.dat};
  \addlegendentry{$\Im\left({\hat{E}_y^{\nu}(x_c)\mathrm{e}^{it}}\right),\; x_c=-2.5125, \; \nu=10^{-3}$};

  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}