### 2015-04-27 report_clean/responce_reviewers.tex
+multiple items changed

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}

%opening
\title{}
\author{}

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Extra Corrections}
As an extra improvement, not asked by the reviewers, we changed $\tilde{\omega}$ to $\omega_{\nu}$ in Section 2.

Corrected the definitions of $\alpha,\delta$ to comply with the definition of the cold plasma dielectric tensor of Stix (more precisely, removed terms $\frac{\omega^2}{c^2}$ in 
front of them)

\section{Responce to the reviewer #1}

All the corrections are marked in blue.

A. General issues.

1. We agree with the remarks of the reviewer. We included a short introduction into the physics of the problem, 
which hopefully will facilitate in reading the article. The Section 4.3.1 contains of course a mistake in the description of the 
parameters of the problem.

Concerning the notation: from now on $\epsilon^{\nu}_{\omega}(\mathbf{r})$ denotes the cold plasma dielectric tensor, similarly we denote its parameters.

We added explicitely dependence on space in parameters that depend on the spatial variables.

Indeed, $\epsilon_{\infty}<\infty$ if $\omega_c^2\neq \omega^2$, but formally one also needs to require the boundedness of $\omega_p^2$. Therefore we 
added an assumption in the end of the Section Frequency Domain Study on the coefficients of the dielectric tensor, and made a remark on the absence
of the cyclotron resonance.

A brief description of what limiting absorption and limiting amplitude principles had been added in the end of the introduction.

2. The reduction to a 1d problem is not clearly justified - indeed, we changed this in the introduction by adding an assumption on $N_e, B_0$, 
and changed the wording in the introduction to underline that the 1D case is the one of interest.

Specific remarks:

- p1, abstract: changed

- p2: in the frequency domain, as we have shown in Lemma 2.1, the problem is regularized by a choice $|\nu|>0$, independently of the sign of $\nu$. 
It is interesting however that for a fixed $\omega$ the limits $\lim_{\nu\rightarrow 0\pm}$ do not result in the same solution. 
In the time domain the sign of $\nu>0$ is indeed of crucial importance for the stability. We added a remark on this before Section 2.1

- p2, after (1): $L$ is just a non-negative real number. We found it more convenient to place the point of the isolated hybrid resonance in $x=0$.
For $\nu\neq 0$ the energy decays, we added this in the introduction.

- p3: yes, indeed, fixed

- p3 eq (4): the notation had been changed

- p4, top: we changed the notation

- p4, eq. (5) fixed

- p4, eq. (8): changed the notation in eq (8), though a more general result in Lemma 2.1 is formulated with $u,v$.

- p4: we changed $\alpha_{\nu}$ to $\beta_{\nu}$ not to confuse with $\alpha$ in the dielectric tensor. The index $\nu$ 
just underlines the dependence of this angle $\beta_{\nu}$ on $\nu$.

- p. 5: indeed, we changed this

- p.6:  changed

-p 6: $Ex$ changed

-p 8: our code, changed

 - p.8: fixed section number

 -

\end{document}