### 2015-05-03 report_clean/intro.tex

\section{Introduction}
Modelling various phenomena in plasmas is of practical importance for developing new sources of energy 
based on plasma fusion, see the ITER project\footnote{www.iter.org}. 
This article concentrates on studying a phenomenon of hybrid resonance \cite{Stix}, 
which is observed in experiments (see \cite{reflectometers_2006, reflectometers_2010, Dumont_2005}) and is described
mathematically as the non-regularity of
the solutions of Maxwell equations in \mrev{magnetized plasmas} \cite{Despres_2014}. 
The energy deposit is resonant and may exceed by far the energy 
exchange which occurs in Landau damping, see \cite{Freidberg_2007}. 
Contrary to the Landau damping, however, 
hybrid resonance appears in a simpler model coupling 
fluid equations with the non electrostatic part of Maxwell equations.

\urev{
We consider the wave propagation in the cold plasma that consists of particles of single species 
under a uniform in time background magnetic field $\mathbf{B}_0=(0,\;0,\; B_0)$. In two dimensions, in the plane 
perpendicular to the direction of the magnetic field, this process can be described by the time-dependent Maxwell system  \cite{Stix}
\begin{align}
\label{eq:main_system}
\begin{split}
-\varepsilon_0 \partial_t E_x +\partial_y H_z = J_x,\\
-\varepsilon_0 \partial_t E_y -\partial_x H_z = J_y,\\
\mu_0 \partial_t H_z + \partial_x E_y-\partial_y E_x= 0
\end{split}
\end{align}
coupled with an equation for the linear electronic current  $\J = eN_e(\mathbf{r}) \ubf$, 
\begin{align}
\begin{split}
\label{eq:electronmove}
m_e \partial_t u_x =\epsilon_0 e E_x+e B_0(\mathbf{r}) u_y -m_e \nu u_x,\\
m_e \partial_t u_y =\epsilon_0 e E_y-e B_0(\mathbf{r}) u_x -m_e \nu u_y,
\end{split}
\end{align}
where \mrev{$\nu\geq 0$} is the friction  between particles, $e<0$ is the charge of electrons, $m_e$ the electron mass and $N_e$ the electron density that in general 
depends on spatial variables and is uniform in time. A non-zero value of $\nu$ constitutes a basis for the limiting absorption principle. 
The unknowns are the electron velocity $\ubf$ and the electromagnetic field.

The energy of this system in a domain $\Omega\in\mathbb R^2$ can be expressed as, see \cite{Despres_2014}
\ben
{\mathcal E}(t)= \int_\Omega \left(
\frac{\epsilon_0 |\E(t,\mathbf{r})|^2}{2}+\frac{ |\B(t,\mathbf{r})|^2}{2\mu_0}+\frac{m_e|\J (t,\x)|^2}{2|e|N_e(\mathbf{r})} 
\right)\mathrm{d}\mathbf{r},
\een
and
\ben
\frac{d}{dt}{\mathcal E}=-\nu\left(\|u_{x}\|_{L_{2}(\Omega)}^2+\|u_{y}\|_{L_{2}(\Omega)}^2\right)+\text{boundary terms}.
\een

To illustrate unexpected difficulties that can occur in this system, let us rewrite it in the frequency domain. To do so, 
we introduce the plasma frequency $\omega_p(\mathbf{r})=\sqrt{\frac{|e|^2N_e(\mathbf{r})}{m\epsilon_0}}$, 
the cyclotron frequency $\omega_c(\mathbf{r})=\frac{e B_0(\mathbf{r})}{m_e}$ and perform the Fourier transform 
in time of (\ref{eq:main_system}) (with the convention $\partial_t\rightarrow -i\omega$), for a 
particular case $\nu=0$:
\begin{align}
\label{eq:model_freq_domain}
 \mathbf{curl}{\operatorname{curl}}\hat{\mathbf{E}}-\frac{\omega^2}{c^2}\underline{\varepsilon}^{0}_{\omega}(\mathbf{r})\hat{\mathbf{E}}=0,
\end{align}
where the cold-plasma dielectric tensor \cite[Chapter 1-2]{Stix}
\begin{align}
\nonumber
\underline{\varepsilon}^{0}_{\omega}(\mathbf{r})=\left(
 \begin{matrix}
  \alpha_{\omega}(\mathbf{r})& i\delta_{\omega}(\mathbf{r})\\
  -i\delta_{\omega}(\mathbf{r}) & \alpha_{\omega}(\mathbf{r})
 \end{matrix}
\right),\\
\label{eq:alpha_delta}
\alpha_{\omega}(\mathbf{r})=1-\frac{\omega_p^2(\mathbf{r})}{\omega^2-\omega_c^2(\mathbf{r})},\; \delta_{\omega}(\mathbf{r})=\frac{\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})}{\omega(\omega^2-\omega_c^2(\mathbf{r}))}.
\end{align}

The point $\mathbf{r}_0$ s.t. $\omega_c(\mathbf{r}_0)=\omega$, i.e. when $\underline{\varepsilon}_{\omega}^{0}(\mathbf{r}_0)$ is singular, is called a cyclotron resonance (c.f. \cite[Chapter 1-5]{Stix}). 
In \cite{singular_solutions} it was demonstrated for a 1D-counterpart of the above system that such points behave like removable singularities, 
hence this case is not of interest for the present work.

Another interesting case is $\omega_p^2(\mathbf{r})+\omega_c^2(\mathbf{r})=\omega^2$ and $\omega_c(\mathbf{r})\omega_p^2(\mathbf{r})\neq 0$, i.e. when the diagonal part of the tensor 
$\underline{\varepsilon}_{\omega}^{0}(\mathbf{r})$ vanishes and the non-diagonal part remains non-zero. This is a hybrid resonance case.

As shown in \cite{Despres_2014, singular_solutions}, for a 1D-counterpart of (\ref{eq:model_freq_domain}), in this case the time-harmonic electric field component $E_x$ is not necessarily square 
integrable. This apparent paradox is of course the source of important numerical difficulties which are the subject of the present study.

Before proceeding, let us define the geometrical setting of the problem. We consider the frequency domain problem (\ref{eq:model_freq_domain}) posed in the domain $(-L,\;\infty)\times\mathbb{R},\; L>0$, with the 
 boundary conditions chosen as follows. At the left boundary of the domain, which represents the wall of the Tokamak, we pose Robin boundary conditions 
\be
\label{eq:boundary_conditions_intro}
curl \hat{\mathbf{E}} -\imath \lambda\hat{\mathbf{E}}\wedge \n = \hat{g}_{inc} =curl \E_{inc} -\imath\lambda\E_{inc}\wedge \n,
\ee
where $\E_{inc} = \exp \left(\imath \lambda x\right)\begin{pmatrix} E_1\\ E_2 \end{pmatrix}$ and $\lambda$ is 
the frequency of the antenna. We truncate the domain 
$(-L,\; \infty)\times\mathbb{R}$ to $(-L,\; H)\times\mathbb{R},\; H>0,$ and set on the right boundary $\operatorname{curl} \hat{\E} = 0$.

In this article we concentrate on a simplified model (which nevertheless is of physical interest when modelling antennas in plasmas), 
where it is required to study the above system under the assumption that the electron density $N_{e}$ and the magnetic field $B_0$ are \textbf{uniform} in the second coordinate, more precisely 
$N_{e}=N_{e}(x)$ and $B_0=B_0(x)$, see e.g. \cite{Despres_2014}. Even this simplified case contains difficulties intrinsic to hybrid resonance \cite{Despres_2014, singular_solutions}.

The one-dimensional model can be derived from (\ref{eq:main_system}) by performing the Fourier transform in $y$ ($\partial_y\rightarrow i\theta, \; \theta\in\mathbb{R}$), 
under the assumption that $N_e=N_e(x), \; B_0=B_0(x)$

\begin{equation}
 \label{eq:main_model}
\begin{cases}
\epsilon_0\partial_t E_{x}-i\theta H_z=-eN_e u_x,\\
\epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\\
\mu_0\partial_t H_z+\partial_x E_y-i\theta E_x =0,\\
m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\\
m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y
\end{cases}
\end{equation}
 posed in $(-L,\; H)$, for some $L,H>0$, with \mrev{the (simplified) boundary conditions compatible with (\ref{eq:boundary_conditions_intro})}
 \begin{align}
 \label{eq:boundary_conditions_frequency_domain}
  \left.\left(-\partial_x E_y -\imath \lambda E_y\right)\right|_{x=-L} =g_{inc},\\
  \left.\partial_xE_y\right|_{x=H}=0.
 \end{align}
This is the time-domain model that will be considered in the rest of the article, unless stated otherwise.

In the frequency domain (\ref{eq:main_model}) for $\nu=0$ is of the form 
\begin{align}
\label{eq:main_frequency_domain_intro}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\underline{\varepsilon}_{\omega}^{0}(x)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
see (\ref{eq:alpha_delta}). Indeed, the coefficients $\alpha_{\omega},\;\delta_{\omega}$ depend solely on $x$ and $\omega$.  
\arev{
We concentrate on a physical situation of the hybrid resonance, which we define as in \cite{singular_solutions}.
\begin{df}
Let $\omega\in\mathbb{R}$ be fixed. 
We will call a point $x_h\in \mathbb{R}$ a hybrid resonance if $\alpha_{\omega}$ has 
a simple root in $x_h$ and has no other roots in an $\epsilon$-vicinity of $x_h$ $B_{\epsilon}(x_h)$, for some $\epsilon>0$, and 
$\delta(x_h)\neq 0$. 
\end{df}
}

To investigate the behaviour of the solutions in the vicinity of the hybrid resonance, 
instead of the above problem we will study a sequence of the regularized problems 
\begin{align}
\label{eq:seq_regularized}
\left(
\begin{matrix}
 i\theta \partial_x \hat{E}_y+\theta^2 \hat{E}_x\\
 i\theta \partial_x \hat{E}_x -\partial_x^2 \hat{E}_y
\end{matrix}
\right)-\frac{\omega^2}{c^2}
\left(\underline{\varepsilon}_{\omega}^{0}(x)+i\nu \mathrm{Id}\right)\left(
\begin{matrix}
 \hat{E}_x\\
 \hat{E}_y
\end{matrix}
\right)
=0,
\end{align}
where $\mathrm{Id}$ is the identity matrix. More details on the connection of this problem with (\ref{eq:main_model}) will be provided in further sections.

 %This domain represents the physical case of a wave sent from a wall facing a plasma.

%In  \cite{Despres_2014} it was shown
%that the time harmonic electric field component $E_x$ in this case is not necessarily square 
%integrable, and  
%explicit estimates on the behavior of the solutions of \eqref{eq:main_model} in 1D were provided.

%some given 
%frequency. 
%We truncate the domain 
%\ben
%-curl \E -\imath \lambda\E \wedge \n = g_{inc} = -curl \E_{inc} -\imath\lambda\E_{inc} \wedge \n,
%\een
%where $\E_{inc} = \exp \left(\imath \lambda x\right)\begin{pmatrix} E_1\\ E_2 \end{pmatrix}$.

%\begin{remark}
%	When modelling antennas in plasma physics \mrev{(e.g. in such applications as ITER)}, 
%	a good choice of boundary condition could be $\lambda = \sqrt{\alpha_{\omega}(-L)}$
%\end{remark}

The goal of this article is three-fold. 
First, we investigate the finite element approximation of the 1D 
problem (\ref{eq:seq_regularized}), with a dielectric tensor \mrev{regularized
with the help of a small parameter $\nu$}. 
%linearized with respect to $\nu$.
We prove the well-posedness of this problem for \mrev{$\nu\neq 0$} in Section~\ref{sec:wellposedness} and 
demonstrate that the use of  (P1) finite elements allows to approximate the singularity 
of the solution fairly well (Section \ref{sec:freq_dep}). 
}

Second, we briefly develop an original scheme based on widely appreciated semi-lagrangian schemes
for the discretization of time domain Maxwell's equations with a linear current.
Thirdly, we consider the case $\nu\rightarrow 0$, and study the limiting amplitude solution 
$\lim\limits_{t\rightarrow +\infty}\lim\limits_{\nu\rightarrow 0}\mathbf{E}(t)$ obtained with the help of 
the FDTD discretization of \eqref{eq:main_model}, suggested in \cite{stable_yee_plasma_current}. 
We compare this result with 
$\hat{\mathbf{E}}\mathrm{e}^{i\omega t}$, computed in the frequency domain, for $\nu\rightarrow 0$, which is equivalent to considering
$\lim\limits_{\nu\rightarrow 0}\lim\limits_{t\rightarrow+\infty}\mathbf{E}(t)$.
\urev{Such a comparison is a way to study the formal commutation relation between the limited absorption and the limiting amplitude principles. 
In application to e.g. non-resonant case of (\ref{eq:model_freq_domain}), the limited absorption principle states that with 
appropriately chosen boundary conditions the solution of (\ref{eq:seq_regularized}) 
as $\nu\rightarrow 0\pm$ 
would coincide with the solution of (\ref{eq:model_freq_domain}). Similarly, the limiting amplitude principle applied to the problem (\ref{eq:main_system}) 
states that if the boundary conditions are chosen harmonic in time, e.g. $\partial_x E_y=\mathrm{e}^{-i\omega_{*} t}$ on 
the boundary, then the solution of this problem $(\mathbf{E}(t),\;H(t))$ asymptotically ($t\rightarrow +\infty$)
tends to a steady state of the form $ \left(\mathbf{E}(\omega_{*}),\; \hat{H}(\omega_{*})\right)\mathrm{e}^{-i\omega_{*} t}$, where 
$\left(\hat{\mathbf{E}}(\omega_{*}),\; \hat{H}(\omega_{*})\right)$ is the solution of (\ref{eq:main_model})
in the frequency domain, with $\omega=\omega_{*}$. For more details see e.g. \cite{Morawetz, Eidus}. The commutation between the two principles then write as}
$$
\lim\limits_{\nu\rightarrow 0}\lim\limits_{t\rightarrow+\infty}= \lim\limits_{t\rightarrow+\infty}\lim\limits_{\nu\rightarrow 0},
$$
see Figure \ref{fig:limits}. 
Even though it is true for standard linear wave problems, it is not clear \mrev{if it still holds} in our case due to the singularity
of the hybrid resonance.
\begin{figure}
\begin{tikzpicture}
   \matrix (m) [matrix of math nodes,row sep=3em,column sep=4em,minimum width=2em] {
     \mathbf{E}^{\nu}(t) & \mathbf{E}^{0+}(t)\\
     \hat{\mathbf{E}}^{\nu}(t)\mathrm{e}^{i\omega t} & 
    \lim\limits_{\nu\rightarrow 0}\lim\limits_{t\rightarrow \infty}\mathbf{E}^{\nu}(t)=\lim\limits_{t\rightarrow \infty}\lim\limits_{\nu\rightarrow 0}\mathbf{E}^{\nu}(t)\\
     };
 \path[-stealth,font=\scriptsize]    (m-1-1) edge node [left] {$t\rightarrow+\infty$} (m-2-1);
  \path[-stealth, font=\scriptsize]   (m-1-1) edge node [above] {$\nu\rightarrow 0+$} (m-1-2);
 \path[-stealth, font=\scriptsize]   (m-2-1) edge node [above] {$\nu\rightarrow 0$} (m-2-2);
  \path[-stealth, font=\scriptsize]   (m-1-2) edge node [right] {$t\rightarrow +\infty$} (m-2-2);

  % \path[-stealth] (m-2-1.east) edge node [above] {$\nu\rightarrow 0$} (m-2-2);
   %(m-1-2) edge node [right] {$\nu\rightarrow 0$} (m-2-2);
\end{tikzpicture}
 \caption{Schematic representation of the equivalence of the limited absorption and limiting amplitude principles.}
\label{fig:limits}
\end{figure}

To our knowledge, such numerical studies have not been performed in the existing literature.