### 2015-05-04 report_clean/figure_generator/pics_time_domain/airy/figure_nu1e2_2.tex
figure update

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=newest}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=\textwidth, 
  height=0.5\textwidth,
  xlabel={$x$},
  enlarge x limits=false,
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ymax=0.8,
  xmin=-0.45,
  ylabel={$E_y^{\nu}(x,t_c)$},
  legend style={draw=none, fill=white, font=\small}
   ]
 \addplot[color=blue, line width=1] file{airy/ey_td_nu1e2_t_29730.dat};
  \addlegendentry{$E_y^{\nu}(x,t_c),\; t_c=29730, \; \nu=10^{-2}$};
  \addplot[color=red, mark=*, dashed, mark size=0.6] file{airy/ey_fd_nu1e2_t_29730.dat};
 \addlegendentry{$\Im\left(\hat{E}_y^{\nu}(x)\mathrm{e}^{-it_c}\right),\; t_c=29730, \; \nu=10^{-2}$};
  \addplot[color=blue, mark=*, dashed, mark size=0.6] file{airy/ey_td_nu1e2_t_29960.dat};
  \addlegendentry{$E_y^{\nu}(x,t_c),\; t_c=29960, \; \nu=10^{-2}$};
  \addplot[color=red] file{airy/ey_fd_nu1e2_t_29960.dat};
  \addlegendentry{$\Im\left(\hat{E}_y^{\nu}(x)\mathrm{e}^{-it_c}\right),\; t_c=29960, \; \nu=10^{-2}$};
  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}