### 2015-06-10 report_clean_final/TEX/abstract.tex

The hybrid resonance is a physical phenomenon
that appears for example in the heating of plasma,
and as such is of scientific interest for the development  of  the ITER project.
In this paper \urev{we focus on} solutions of low regularity to Maxwell equations in
\mrev{magnetized plasmas}. 
Our main purpose is three-fold. First, we aim at investigating the finite element approximation 
of the frequency-domain problem. Second, we would like to study the resonant solutions 
in the time domain, with the help of two different finite difference approximations. 
We finally  compare the results with the ones obtained in the frequency domain, by  
numerical examination of the limiting absorption and limited amplitude principles.