### 2015-06-10 report_clean_final/TEX/figure_generator/pics_frequency_domain/figure_cond_number.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[frenchb,english]{babel}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}

\usepackage{color}
\newenvironment{remark}[1][Remark]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{red}{rgb}{1,0.3,0.3}
\usetikzlibrary{positioning,shapes,shadows,arrows}
\begin{document}
\pagenumbering{gobble}
\begin{figure}
\begin{tikzpicture}
    \begin{loglogaxis}[
    	width=\textwidth,
    	height=\textwidth,
        xlabel=$\nu$,
        ylabel={Condition number},
        ymin=1,
        legend style={
at={(0.2,1.5)},
legend pos=north  east,
%anchor=north east, 
draw=none, 
%font=\tiny
}
]
\addplot[blue, mark=*] table {cond_number_pic/g_h1.dat}; 
\addlegendentry{$h=1$};
\addplot[black, mark=triangle] table {cond_number_pic/g_h0.5.dat}; 
\addlegendentry{$h=0.5$};
\addplot[red, mark=square] table {cond_number_pic/g_h0.1.dat}; 
\addlegendentry{$h=0.1$};
\addplot[black, mark=*] table {cond_number_pic/g_h0.01.dat}; 
\addlegendentry{$h=0.01$};
\addplot[blue,dashed] table {cond_number_pic/1nu.dat};
\addlegendentry{$O(\nu^{-1})$};
 \end{loglogaxis}
\end{tikzpicture}
\end{figure}

\end{document}