### 2015-06-10 report_clean_final/TEX/figure_generator/pics_time_domain/airy/figure_nu1e2.tex

\documentclass[a4paper,10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array}
\usepackage{tikz}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\pgfplotsset{compat=1.10}
\usepackage{hyperref}

\begin{document}
\pagenumbering{gobble}% Remove page numbers (and reset to 1)
\clearpage
\thispagestyle{empty}
\begin{figure}
\begin{tikzpicture}
 \begin{axis}[
  width=1.2\textwidth, 
  height=0.5\textwidth,
   xlabel={$t$},
  enlarge x limits=false,
  ymax=0.5,
 % xmin=2.97e4,
 % xtick={2.97e4,2.98e4,2.99e4,3e4},
  x tick label style={/pgf/number format/fixed, /pgf/number format/precision=4},
  ylabel={$E_y^{\nu}(x_c,t)$},
  legend style={draw=none, fill=white, font=\tiny}
   ]
 \addplot[color=blue] file{airy/sol_1e2_dt0.05_timedomain.dat};
  \addlegendentry{$E_y^{\nu}(x_c,t),\; x_c=0.45, \; \nu=10^{-2}$};
  \addplot[color=red, mark=*, dashed, mark size=0.6] file{airy/sol_1e2_dt0.05_freqdomain.dat};
  \addlegendentry{$\Im\left(\hat{E}_y^{\nu}(x_c)\mathrm{e}^{-it}\right),\; x_c=0.45, \; \nu=10^{-2}$};
  \end{axis}
\end{tikzpicture}
\end{figure}
\end{document}