### 2015-06-10 report_clean_final/TEX/proc.tex

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a template file for Proccedings 
%
% Copy it to a new file with a new name and use it as the basis
% for your article
%
%%%%%%%%%%%%%%%%%%%%%%%%   EDP Sciences  %%%%%%%%%%%%%%%%%%%%%%%%%%
%
\documentclass[proc]{edpsmath}
%
%\usepackage{amsfonts,graphicx,amssymb}%amsfonts pour les \mathbb
%\usepackage{amsmath,amsthm,amssymb}
%%%%%%%%%%%%%--PREAMBLE--%%%%%%%%%%%%%%%%%%
%%-----------------------------
%%         ...........
%%         your macros
\usepackage[frenchb,english]{babel}
\usepackage{macros}
\usepackage{tikz}
\usepackage{xcolor}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amsfonts,amssymb,amsthm,epsfig,epstopdf,array, ulem}
\usepackage{tikz}
\usetikzlibrary{matrix,arrows,decorations.pathmorphing}
\usepackage{caption}
\usepackage{pgfplots}
\usepackage{ulem}
\usepackage{pifont}% http://ctan.org/pkg/pifont
\usepackage{hyperref}
\usepackage{diagbox}
\usepackage{color}
\usepackage{placeins}
%\newenvironment{remark}[1][Remark]{\begin{trivlist}
%\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{df}{Definition}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{assumption}[theorem]{Assumption}
\definecolor{blue}{rgb}{0.30,0.8,1}
\definecolor{greem}{rgb}{0,0.5,0}
\definecolor{bluer}{rgb}{0,0,1}
\definecolor{red}{rgb}{1,0.3,0.3}
\DeclareMathOperator{\Ai}{Ai}
%\usetikzlibrary{positioning,shapes,shadows,arrows}

%%         ...........
%%-------------------------%%----
%%%%%%%%%%%%%%%--BODY--%%%%%%%%%%%%%%%%%%

\begin{document}
%\newcommand{\urev}[1]{%{\color{bluer}#1}   
\newcommand{\urev}[1]{#1}
%\newcommand{\arev}[1]{{\color{red}#1}}
\newcommand{\arev}[1]{#1}
%\newcommand{\mrev}[1]{{\color{greem}#1}}
\newcommand{\mrev}[1]{#1}
%%color printing
%\newcommand{\urev}[1]{{\color{black}#1}}
%\newcommand{\arev}[1]{{\color{black}#1}}
%\newcommand{\mrev}[1]{{\color{black}#1}}

%%-----------------------------
%%      the top matter
%%-----------------------------
\title{A numerical study of the
	solution of X-Mode equations around the hybrid resonance}
\thanks{\rm The authors acknowledge 
	the support
	of ANR under contract ANR-12-BS01-0006-01. Moreover,  
	this work was carried out within the framework of the European Fusion Development Agreement and 
	the French Research Federation for Fusion Studies. 
	It is supported by the European Communities under the contract of Association between Euratom and CEA. 
	The views and opinions expressed herein do not necessarily reflect those of the European Commission.} %\thanks{...}\thanks{...}% At most 5 thanks
%
\author{C\'eline Caldini-Queiros}\address{Max Planck Institute für PlasmaPhysik, Garching bei München, Germany}
\author{Bruno Despr\'es}\address{Sorbonne Universit\'es, UPMC Univ Paris 06, UMR 7598, Laboratoire Jacques-Louis Lions, F-75005, Paris, France}
\author{Lise-Marie Imbert-Gérard}\address{Courant Institute of Mathematical Sciences, New York University}
\author{Maryna Kachanovska}\address{POEMS, INRIA, ENSTA ParisTech, Palaiseau, France}

%
%\dedicated{\it Dedicated to Maurice Dupont} %if necessary
%
\begin{abstract}
\input{abstract}
\end{abstract}
%
\begin{resume} 
\input{resume}
\end{resume}
%
%
\maketitle

\input{intro}
\input{FD}%\input{equations}

\input{timedomain}

\input{numerics}

\input{absorption}

\input{conclusions}

\begin{thebibliography}{99}

\bibitem{abramowitz_stegun} Abramowitz, M. and Stegun, I. A., eds. (1972). Handbook of Mathematical Functions with Formulas, Graphs, and Mathematical Tables. New York: Dover Publications.
	Handbook of mathematical functions
\bibitem{brenner} Brenner, S. C. and Scott, R. (2008). The mathematical theory of finite element methods (Vol. 15). Springer.

\bibitem{cedar}
Charles, F., Despr\'es, B., Mehrenberg, M.  (2013).  
Enhanced convergence estimates for semi-lagrangian schemes Application to the Vlasov-Poisson equation, SIAM J. Numer. Anal., 51(2), 840-863.

\bibitem{stable_yee_plasma_current} Da Silva, F., Campos-Pinto, M., Després, B., and Heuraux, S. (2014). Stable coupling of the Yee scheme with a linear current model.

\bibitem{compfluids}
Del Pino, S., Despr\'es, B., Hav\'e, P., Jourdren, H., and Piserchia, P. F. (2009). 3D finite
volume simulation of acoustic waves in the earth atmosphere.
Comput. \&
Fluids, 38(4):765-777.

\bibitem{reflectometers_2006} Da Silva, F., Heuraux, S and  Manso, M. (2006). Developments on reflectometry simulations for fusion plasmas: application to ITER position reflectometers. Journal of plasma physics, 72(06), 1205-1208.

\bibitem{reflectometers_2010}Da Silva, F., Heuraux, S., Gusakov, E. Z., and Popov, A. (2010). A Numerical Study of Forward-and Backscattering Signatures on Doppler-Reflectometry Signals. Plasma Science, IEEE Transactions on, 38(9), 2144-2149.

\bibitem{Despres_2014} Després, B., Imbert-Gérard, L. M. and Weder, R. (2014). Hybrid resonance of Maxwell's equations in slab geometry. Journal de Mathématiques Pures et Appliquées, 101(5), 623-659.

\bibitem{Dumont_2005} Dumont, R. J., Phillips, C. K., and Smithe, D. N. (2005). Effects of non-Maxwellian species on ion cyclotron waves propagation and absorption in magnetically confined plasmas. Physics of Plasmas (1994-present), 12(4), 042508.

\bibitem{feix} Feix, M. R., Bertrand, P., and Ghizzo, A.  (1994). Eulerian codes for the Vlasov equation.  Series on Advances
in Mathematics for Applied Sciences 22, Kinetic Theory and Computing, pp. 45.

\bibitem{fifi} 
 Filbet, F., Sonnendrucker, E.  (2003). 
 Comparison of Eulerian Vlasov solvers,
Computer Physics Communications,
150, pp. 247-266

\bibitem{Freidberg_2007} Freidberg, J. P. (2007). Plasma physics and fusion energy. Cambridge University Press.

\bibitem{guo}
Guo, W. and Qiu, J. - M. (2013). 
Hybrid semi-Lagrangian finite element-finite difference methods for the Vlasov
equation,
Journal of Computational Physics,
234, pp. 108-132.

\bibitem{phdhattori}
Hattori, T. (2014).
D\'ecomposition de domaine pour la simulation Full-Wave dans un plasma froid.
Ph.D. Thesis, Universit\'e de
Lorraine.

\bibitem{LMIG_thesis} Imbert-Gérard, L. M. (2013). Analyse mathématique et numérique de problèmes d'ondes apparaissant dans les plasmas magnétiques (Doctoral dissertation, Université Pierre et Marie Curie-Paris VI).

\bibitem{simon}
Labrunie, S., Bertrand, P., Roche, J.-R., Back, A., and
Hattori, T. 
Electromagnetic wave propagation and absorption in
magnetised plasmas: variational formulations and
domain decomposition, Hal preprint 2014, https://hal.archives-ouvertes.fr/hal-01075137.

\bibitem{Morawetz} Morawetz, C. S. (1962). The limiting amplitude principle. Communications on Pure and Applied Mathematics, 15(3), 349-361.

\bibitem{Eidus} Eidus, D.M. (1969). The principle of limit amplitude. Russian Mathematical Surveys, vol. 24, issue 3, 97-167.

\bibitem{soso}
Sonnendrucker, E., Roche, J., Bertrand, P., and A. Ghizzo (1998). 
The Semi-Lagrangian Method for the
Numerical Resolution of Vlasov Equations.
J. Comput. Phys.,
149, pp. 201-220.

\bibitem{Stix} Stix, T. H. (1992). Waves in plasmas. Springer.
\bibitem{Xu_2006}Xu, L., and Yuan, N. (2006). FDTD formulations for scattering from 3-D anisotropic magnetized plasma objects. Antennas and Wireless Propagation Letters, IEEE, 5(1), 335-338.

\bibitem{umeda}
Umeda, T., Nariyuki, Y. and Kariya, D. (2012).
A non-oscillatory and conservative semi-Lagrangian scheme with
fourth-degree polynomial interpolation for solving the Vlasov equation,
Computer Physics Communications,183, pp. 1094-1100.

\bibitem{singular_solutions}
Despr\'es, B., Imbert-G\'erard, L.-M. and Lafitte, O., 2014. Singular solutions for the plasma at the resonance. $\langle$ hal-01097364$\rangle$

\bibitem{ciarlet_fem}
Ciarlet, P. G. (1978).  The Finite Element Method For Elliptic Problems. North-Holland Publishing Company.
\end{thebibliography}

\end{document}