### 2015-09-14 report_clean_final/TEX/timedomain.tex
Correcting typoes as asked by the referee

\section{Time-Dependent Problem Discretization}
\arev{
We develop hereafter the discretization of the time-dependent one-dimensional model (\ref{eq:main_model}), for the particular case $\theta=0$. 
We first show how this system can be discretized thanks to a Yee scheme suggested in \cite{stable_yee_plasma_current}, then a scheme based on semi-lagrangian discretization is employed.}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{A staggered scheme}
\label{sec:staggered_discretization}
%In this section, we develop a discretization of the time-dependent problem~\eqref{eq:main_model} in one dimension. %In 1D, the derivatives with respect to $y$ vanish, and we are left with the following problem, 
% \begin{align}
% \epsilon_0\partial_t E_{x}=-eN_e u_x, \nonumber\\
% \epsilon_0\partial_t E_{y}+\partial_x H_z=-eN_e u_y,\nonumber\\
% \mu_0\partial_t H_z+\partial_x E_y-\partial_y E_x=0, \label{eq:tdpb}\\
% m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\nonumber\\
% m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.\nonumber
% \end{align}
We compute a numerical approximation of~\eqref{eq:main_model} \mrev{with $\theta=0$} on a cartesian grid using the Yee scheme, 
as suggested in \cite{stable_yee_plasma_current}. Let $n$ and $\Delta t$ denote the time step index and the time step, as usual the discrete time is $t_n = n \Delta t$. 
It is standard (see \cite{Xu_2006,stable_yee_plasma_current }) to discretize the electrical field on the integer time steps, 
whereas the field $H_z$ on half time steps, $t_{n+1/2}$. The time discretization of the system reads:
 \bealn
 \left\{\begin{array}{l}
 \epsilon_0\frac{ E_{x}^{n+1}-E_x^{n}}{\Delta t}=-eN_e \frac{u_x^{n+1}+u_x^{n}}{2}, \nonumber\\
 \epsilon_0\frac{ E_{y}^{n+1}-E_y^{n}}{\Delta t}+\partial_x H_z^{n+1/2}=-eN_e \frac{u_y^{n+1}+u_{y}^{n}}{2},\nonumber\\
 \lastrev{\mu_0\frac{ H_z^{n+1/2}-H_z^{n-1/2}}{\Delta t}+\partial_x E_y^{n}=0,} \label{eq:tdiscretizedpb}\\
 m_e\frac{ u_{x}^{n+1}-u_x^{n}}{\Delta t}=e\frac{E_x^{n+1}+E_x^{n}}{2}+e\frac{u_y^{n+1}+u_{y}^{n}}{2}B_0-\nu m_e \frac{u_x^{n+1}+u_x^{n}}{2},\nonumber\\
 m_e\frac{u_{y}^{n+1}-u_y^{n}}{\Delta t}=e\frac{E_y^{n+1}+E_y^{n}}{2}-e\frac{u_x^{n+1}+u_x^{n}}{2}B_0-\nu m_e \frac{u_y^{n+1}+u_y^{n}}{2}.\nonumber
\end{array}\right.
 \eealn
In \lastrev{the} Yee scheme, the different components of the vector fields 
are defined on staggered points. 
We summarize this discretization on Fig.~\ref{schemepos} where the positions 
of \urev{$E_x$}, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) 
on the discretized grid in space and time are displayed. The fully discretized system then reads

\ben 
\eps_0 \frac{E_x\inp- E_x\midin}{\Delta t} = -e N_e \frac{u_x\inp + u_x \midin}{2},
\label{eq:ns1}
\een 
\ben
\eps_0 \frac{E_y\ihnp-E_y\ihn}{\Delta t}+ \frac{H_z\ipnh-H_z\inh}{\Delta x} = -e\lastrev{N_e} \frac{u_y\ihnp + u_y\ihn}{2} ,
\label{eq:ns2}
\een
\ben
\lastrev{\mu_0} \frac{H_z\inh-H_z\inmh}{\Delta t}+ \frac{E_y\ihn-E_y\imhn}{\Delta x}=0,
 \label{eq:ns3}
\een
\ben 
m_e \frac{u_x\inp-u_x\midin }{\lastrev{\Delta t}} = \lastrev{e} \frac{E_x\inp+E_x\midin}{2}-\nu m_e\frac{u_x\inp+u_x \midin}{2}+ eB_0\frac{u_y\ihn+u_y\ihnp}{2},
\label{eq:ns4}
\een
\ben
m_e \frac{u_y\ihnp-u_y\ihn}{\Delta t} = \lastrev{e}\frac{E_y\ihnp + E_y\ihn}{2}- \nu m_e \frac{u_y\ihnp +u_y\ihn}{2}-eB_0 \frac{u_x\midin+u_x\inp}{2},
\label{eq:ns401}
\een
%\input{grid}
\begin{figure}[h!]
	\begin{center}
 \scalebox{1.2}{
		\begin{tikzpicture}

		\draw[thin] (0, 0) grid (2, 2);
		\foreach \x in {0,1,2}
		\foreach \y in {0,1,2}
		\draw[fill](\x,\y) circle (2pt) ;
		\foreach \x in {0.5,1.5}
		\foreach \y in {0,1,2}
		\draw[fill,blue](\x,\y) circle (2pt) ;
		\foreach \x in {0,1,2}
		\foreach \y in {0.5,1.5}
		\draw[fill,red](\x,\y) circle (2pt) ;		
		\draw[->] (0,0) -- (0,2.5);
		\draw[->] (0,0) -- (2.5,0);
		\coordinate (Origin)   at (0,0);
		\node (xlab) at (2.5,-0.5){\small $x$};
		\node (tlab) at (-0.5,2.5){\small $t$};
		%      \node[black] (exux) at (2.6,1){\tiny $E_x$, $u_x$};
		%      \node[blue](eyuy) at (1.5,2.5){\tiny $E_y$, $u_y$};
		%      \node[red] (hz) at (2.5,1.5) {\tiny $H_z$};
		\node[black] (xilab) at (-0.5,0){\tiny $(x_i, t^n)$};
		\node[blue](xid) at (0.5,-0.3){\tiny $(x_{i+1/2},t^n)$};
		\node[red] (tid) at (-0.8,0.5) {\tiny $(x_i,t^{n+1/2})$};

		\end{tikzpicture} 
}
	\end{center}
	\caption{Positions of $E_x$, $u_x$ (black dots), $E_y$, $u_y$ (blue dots) and $H_z$ (red dots) on the discretized grid in space and time}

	\label{schemepos}
\end{figure}

It is straightforward to rewrite this system in an explicit form (we omit here these simple but slightly tedious computations).
%It is now possible to describe explicitly how to solve the proviso discrete system. We first compute $H_z$ at time $n+1/2$ with \eqref{eq:ns3}. Then, \eqref{eq:ns2} gives the expression $E_y\ihnp$ as a function of known variables, and $u_y^{n+1}$, as follow
%\be
%E_y\ihnp = E_y\ihn - \frac{\Delta t }{\eps_0}\frac{H_z\ipnh-H_z\inh}{\Delta x} - \frac{\Delta t e N_e}{\eps_0}\frac{u_y\ihnp + u_y\ihn}{2}.
%\label{eq:ns5}
%\ee
%
%By incorporating this expression into \eqref{eq:ns401} we then obtain $u_y$ at time $n+1$ as a function of known variables and $u_x^{n+1}$, 
%\be 
%\begin{array}{l}
%u_y\ihnp m_e = m_e u_y\ihn + \Delta t eN_e \frac{E_y\ihn}{2}\\
%+ \Delta teN_e\left(E_y\ihn - \frac{\Delta t }{\eps_0}\frac{H_z\ipnh-H_z\inh}{\Delta x} - \frac{\Delta t e N_e}{\eps_0}\frac{u_y\ihnp + u_y\ihn}{2}\right) \\
%-\nu \Delta t m_e \frac{u_y\ihnp +u_y\ihn}{2}-\Delta teB_0 \frac{u_x\midin-u_x\inp}{2},
%\end{array}
%\label{eq:ns6}
%\ee
%or equivalently the explicit expression
%\begin{align}
%u_y\ihnp \left(m_e(1+\frac{\Delta t \nu}{2}) +\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}\right) =&\left(m_e(1-\frac{\Delta t \nu}{2}) -\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}\right)u_y\ihn +\Delta t eN_e E_y\ihn \nonumber\\
%&- \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x} - \frac{e Bo \Delta t}{2}\left(u_x\midin+u_x\inp\right).
%\label{eq:ns7} 
%\end{align}
%To simplify this expression, consider
%\begin{align}
%u_y\ihnp K_1&= K_2 u_y\ihn +\Delta t eN_e E_y\ihn-  \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x}- \frac{e Bo \Delta t}{2}\left(u_x\midin+u_x\inp\right).
%\label{eq:ns7bis} 
%\end{align}
%with $K_1 =  m_e(1+\frac{\Delta t \nu}{2}) +\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}$ and $K_2 =m_e(1- \frac{\Delta t \nu}{2}) -\frac{\left(\Delta t e N_e\right)^ 2}{2\eps_0}$.
%
%
%At this point only $E_x^{n+1}$ and $u_x^{n+1}$ still need to be computed. 
%Since \eqref{eq:ns1} reads
%\be 
%E_x\inp = E_x\midin - \frac{\Delta t e N_e}{\eps_0}\frac{u_x\inp + u_x\midin}{2},
%\label{eq:ns8}
%\ee
%we then have the scheme 
%\begin{align}
%m_e u_x\inp= & m_e u_x\midin  + \Delta t e N_e \frac{1}{2}\left( E_x\midin - \frac{\Delta t e N_e}{\eps_0}\frac{u_x\inp + u_x\midin}{2}\right) + \Delta t  e N_e \frac{1}{2}E_x\midin -\Delta t\nu m_e\frac{u_x\inp+u_x \midin}{2}\nonumber\\
%&+  \frac{e B_0 \Delta t}{2K_1}\Big(K_2 u_y\ihn +\Delta t eN_e E_y\ihn -  \frac{(\Delta t )^2 e N_e}{2\eps_0} \frac{H_z \ipnh- H_z \inh}{\Delta x}\nonumber\\
%&- \frac{e B_0 \Delta t}{2}\left(u_x\midin+u_x\inp\right)\Big)+  \frac{1}{2}\Delta teB_0u_y\ihn,
%\label{eq:ns9}
%\end{align}
%which can be simplified as
%\begin{align}
% \left(me(1+\frac{\Delta t \nu}{2})+ \frac{(eB_0\Delta t)^2}{4K_1} + \frac{(e\Delta t N_e)^2}{4\eps_0}\right) u_x\inp 
%= \left(me(1-\frac{\Delta t \nu}{2})- \frac{(eB_0\Delta t)^2}{4K_1} -\frac{(e\Delta t N_e)^2}{4\eps_0}\right) u_x\midin \nonumber\\
% + e \Delta t N_e E_x \midin  + \left( \frac{eB_0 \Delta t K_2}{2 K_1}+ \frac{e B_0 \Delta t}{2}\right) u_y\ihn - \frac{(e\Delta t)^2 N_e B_0}{2K_1}E_y\ihn  - \frac{e^2(\Delta t)^3 B_0}{4K_1}\frac{H_z \ipnh- H_z \inh}{\Delta x},
%\labelnd{align}{eq:ns10} 
%\end{align}
%and the scheme is then fully explicit.
%
%To summarize, the algorithm to solve the discrete system \eqref{} is
%\begin{itemize}
%\item compute $H_z^{n+1/2}$ thanks to \eqref{eq:ns3},
%\item compute $E_y^{n+1/2}$ thanks to \eqref{eq:ns6},
%\item compute $u_y^{n+1}$ thanks to \eqref{eq:ns7bis},
%\item compute $u_x^{n+1}$ thanks to \eqref{eq:ns10},
%\item compute $E_x^{n+1}$ thanks to \eqref{eq:ns8}.
%\end{itemize}
\subsection{A co-localized scheme based on semi-lagrangian discretisation}

The scheme described hereafter is mentioned since it is based on some well-known 
numerical methods in plasma physics, the so-called semi-lagrangian schemes widely used for the numerical approximation of transport equations.
It seems to be an original use of semi-lagrangian schemes 
and is mentioned as possible other option. 
We restrict the presentation to the minimum and leave the details  of the construction to the alert reader.
Concerning the   vast literature on the use of semi-lagrangian schemes
in numerical plasma physics, we refer the reader to \cite{feix,guo,umeda,fifi,soso} and references therein.
	The structure of the following  algorithm is very similar to the one developed in \cite{compfluids} and can be adapted and studied numerically
	in any dimension on a Cartesian grid and for more complex right-hand sides.
	So mathematical properties of the  first step of the following semi-lagrangian scheme have been described in  \cite{cedar}.

The idea is to split 
System (\ref{eq:main_model})
into a series of  transport equations (here just 2) and a local in-the-cell system.

- At time step $t_n=n\Delta t$, all quantities are given $(E_x,E_y, H_z, u_x,u_y)_{i}^n$ in the cells, that is for all $i$ in the 
computational domain.

- The first step is to solve 
$$
\left\{
\begin{array}{lll}
\epsilon_0\partial_t E_{x}=0,\\
\epsilon_0\partial_t E_{y}+\partial_x H_z=0,\\
\mu_0\partial_t H_z+\partial_x E_y=0,\\
m_e\partial_t u_x=0,\\
m_e\partial_t u_y=0.
\end{array}
\right.
$$
Since  $\epsilon_0 \mu_0c^2=1$, this can be  performed solving
two transport equations 
$$
\left\{
\begin{array}{lll}
\partial_t w +c \partial_x w=0, & w=E_{y}+ c H_z,  \\
\partial_t z -c \partial_x z=0& z=E_{y}-  c H_z.
\end{array}
\right.
$$
In our case we used 
a semi-lagrangian scheme for a time step $\Delta t$.
In the tests, we used Strang's stencils described in \cite{cedar} with CFL $\nu=0.5$, at order 7. 
It  yields to the chain 
$(E_y,H_z)_i^n\mapsto ( w,z)_i^n\mapsto ( w,z)_i^{n*}\mapsto (E_y,H_z)_i^{n*} $.
Here $n*$ denotes the intermediate time step of the algorithm.

- The second (and final) step is to solve
$$
\left\{\begin{array}{l}
\epsilon_0\partial_t E_{x}=-eN_e u_x,\\
\epsilon_0\partial_t E_{y}=-eN_e u_y,\\
\mu_0\partial_t H_z=0,\\
m_e\partial_t u_x=eE_x+eu_yB_0-\nu m_e u_x,\\
m_e\partial_t u_y=eE_y-eu_xB_0-\nu m_e u_y.
\end{array}
\right.$$
This is performed locally in every cell starting from the data at the preceding 
step. In our case we just adapt the algorithm developed for the staggered scheme.
We make sure that the method is conservative in total energy if $\nu=0$, using a half-step discretization
of the right-hand sides.

- Therefore  all quantities are now given $(E_x,E_y, H_z, u_x,u_y)_{i}^{n+1}$ in the cells 
at time step $t_{n+1}=(n+1)\Delta t$, and that defines totally a time step of the algorithm.